Do we need detailed demographic data to forecast plant population responses to climate change?

### 2016-05-23 manuscript/manuscript.Rmd
adds PBA edits to ms

**1.** Rapid environmental change has generated growing interest in forecasts of future population trajectories.
Traditional population models, typically built using detailed demographic observations from one study site, can address the impacts of environmental change at one location, but are difficult to scale up to the landscape and regional scales relevant to management decisions.

**2.** An alternative is to build models using population-level data that are much easier to collect over broad spatial scales than individual-level data.
However, it remains unclear if models built using aggregated individual-level data adequately capture the effects of density-dependence and environmental forcing that are necessary to generate skillful forecasts.

**3.** Here, we test the consequences of aggregating individual responses when forecasting the population states and trajectories of four perennial grass species in a semi-arid grassland in Montana, USA.
We parameterized two population models for each species, one based on individual-level data (survival, growth and recruitment) and one on population-level data (percent cover), and compared their forecasting skill and forecast horizons with and without the inclusion of climate covariates.
For both models we used Bayesian ridge regression to identify the optimal predictive model in terms of climate covariate strengths.

**4.** Without climate effects included, we found no significant difference between the forecasting skill of models based on individual-level data and models based on population-level data.
Climate effects were weak and caused only slight increases in forecasting skill.
Increases in skill were similar between model types except for one species where forecast accuracy from the individual-level model was significantly higher than the accuracy from an equivalent population-level model.

**5.** *Synthesis.* For our focal species at this particular location, and using our particular statistical models, demographic data was generally unnecessary to achieve skillful forecasts, though for certain species forecast skill can be gained by using demographic data linked to climate covariates.
We conclude that models based on aggregated individual-level data offer a practical alternative to data-intensive demographic models when species do not respond strongly to interannual variation in weather, but when modeling species that do respond to climate drivers, demographically-based models can generate more skillful forecasts.

@@

Perhaps the greatest challenge for ecology in the 21st century is to forecast the impacts of environmental change [@Clark2001; @Petchey2015]. Forecasts require sophisticated modeling approaches that fully account for uncertainty and variability in both ecological process and model parameters [@Luo2011a; but see @Perretti2013a for an argument against modeling the ecological process]. The increasing statistical sophistication of population models [@Rees2009] makes them promising tools for predicting the impacts of environmental change on species persistence and abundance. But reconciling the scales at which population models are parameterized and the scales at which environmental changes play out remains a challenge [@Clark2010; @Freckleton2011; @Queenborough2011; @Clark2012]. The problem is that most population models are built using data from a single study site because collecting those data, which involves tracking the fates of individuals plants, is so difficult. The resulting models cannot be applied to the landscape and regional scales relevant to decision-making without information about how the fitted parameters respond to spatial variation in biotic and abiotic drivers [@Sæther2007]. The limited spatial extent of individual-level demographic datasets constrains our ability to use population models to address applied questions about the consequences of climate change.

<!--
The inability of most population models to address landscape-scale problems may explain why land managers and conservation planners have embraced species distribution models (SDMs) [see @Guisan2005 for a review]. SDMs typically rely on easy-to-collect presence/absence data [but see @Clark2014 for new methods] and remotely-sensed environmental covariates that allow researchers to model large spatial extents [e.g., @Maiorano2013]. Thus, it is relatively straightforward to parameterize and project SDMs over landscapes and regions. However, the limitations of SDMs are well known [@Pearson2003; @Elith2009; @Araujo2012]. Ideally, researchers would provide managers with landscape-scale population models, combining the extent of SDMs with information about dynamics and species abundances [@Schurr2012; @Merow2014].  
-->

Aggregate measures of population status, rather than individual performance, offer an intriguing alternative for modeling populations [@Clark2004; @Freckleton2011]. Population-level data cannot provide inference about demographic mechanisms, but might be sufficient for modeling future population states, especially since such data are feasible to collect across broad spatial extents [e.g., @Queenborough2011]. The choice between individual and population-level data involves a difficult trade-off: while individual-level data leads to more mechanistic models, population-level data can lead to models that can be applied over greater spatial and temporal extents because the data are easier to collect over large spatial scales. An open question is how much forecasting skill is lost when we build models based on population rather than individual-level data.

To date, most empirical population modelers have relied on individual-level data, with few attempts to capitalize on population-level measures. An important exception was an effort by @Taylor2004 to model the population growth rate of an invasive species to identify the best strategies for invasion control. They used a "density-structured" model where the state variable is a discrete density state rather than a continuous density measure. Such models do not require individual-level demographic data and can adequately describe population dynamics. Building on @Taylor2004, @Freckleton2011 showed that density-structured models compare well to continuous models in theory, and @Queenborough2011 provide empirical evidence that density-structured models are capable of reproducing population dynamics at landscape spatial scales, even if some precision is lost when compared to fully continuous models. However, previous tests of density-structured models have yet to assess their ability to forecast out-of-sample observations, and they have not included environmental covariates, which is necessary to make forecasts of population responses to climate change.

Addressing climate change questions with models fit to population-level data is potentially problematic. It is individuals, not populations, that respond to climate variables [@Clark2011b]. Ignoring this fact amounts to an "ecological fallacy", where inference about the individual relies on statistical inference on the group [@PIANTADOSI1988]. Population growth (or decline) is the outcome of demographic processes such as survival, growth, and recruitment that occur at the level of individual plants. Climate can affect each demographic process in unique, potentially opossing, ways [@Dalgleish2011]. These unique climate responses may be difficult to resolve in statistical models based on population-level data where demographic processes are not identifiable. Futhermore, models based on aggregated data may reflect short-term effects of one vital rate (e.g., plant growth) more than others, but over the long-term the influence of other vital rates may be important. If important climate effects are missed because of the aggregation inherent in in population-level data, then population models built with such data will make uninformative or unreliable forecasts.

Here, we compare the forecasting skill of statistical and population models based on aggregated, population-level data with models based on individual-level data. We used a demographic dataset that tracks the fates of individual plants from four species over 14 years to build two kinds of single-species population models, traditional models using individual growth, survival, and recruitment data and alternative models based on basal cover. We use the models to answer two questions: (1) Can population models fit using aggregated individual-level data (percent cover) adequately capture density dependence to produce forecasts as skillful as those from models fit to demographic data? 
And (2) Can population models fit using aggregated data adequately capture the influence of climate on population growth and, in turn, produce forecasts as skillful as those from models fit to demographic data? 
Both questions are motivated by the fact that the effects of intraspecific competition (density dependence) and interannual weather variability act at the level of the individual [@Clark2011b], meaning that models based on individual-level processes should generate more skillful forecasts.

Materials and Methods
---------------------
### Study site and data
Our demographic data come from the Fort Keogh Livestock and Range Research Laboratory in eastern Montana's northern mixed prairie near Miles City, Montana, USA ($46^{\circ}$ 19' N, $105^{\circ}$ 48' W). The dataset is freely available on Ecological Archives[^link] [@Anderson2011], and interested readers should refer to the metadata for a complete description. The site is about 800 m above sea level and mean annual precipitation (1878-2009) is 334 mm, with most annual precipitation falling from April through September. The community is grass-dominated, and we focused on the four most abundant grass species: \emph{Bouteloua gracilis} (BOGR), \emph{Hesperostipa comata} (HECO), \emph{Pascopyrum smithii} (PASM), and \emph{Poa secunda} (POSE) (Fig. 1). *B. gracilis* is a warm-season perennial grass, whereas *H. comata*, *P. smithii*, and *P. secunda* are cool-season perennial grasses. All species typically begin growth in the early spring, reach maximum growth and flower in early to mid summer (May-June), and disperse seed in mid to late summer (July-September).

@@

We modeled each grass population based on two levels of data: individual and quadrat. The individual data is the "raw" data. For the quadrat-level we data we simply sum individual basal cover for each quadrat by species. This is equivalent to a near-perfect census of quadrat percent cover because previous analysis shows that measurement error at the individual-level is small [@Chu2014]. Based on these two datasets of 13 year-to-year transitions, we can compare population models built using individual-level data and aggregated, quadrat-level data. At the individual level we explicitly model three vital rates: growth, survival, and recruitment. At the quadrat level we model population growth as change in percent cover of quadrats with non-zero cover in year *t* and in year *t-1*. For modeling population growth at the quadrat level we ignore within-quadrat extirpation and colonization events because they are very rare in our time series ($N=16$ and $N=13$, respectively, across all species). Sample sizes for each species and vital rate model are shown in Table 1. Given the relatively broad spatial distribution of the quadrats we are studying, it is safe to assume that these events are in fact rare enough to be ignored for our purposes.

@@

### Stastical models of vital rates
At both levels of inference (individual and quadrat), the building blocks of our population models are vital rate regressions. For individual-level data, we fit regressions for survival, growth, and recruitment for each species. At the quadrat-level, we fit a single regression model for population growth. We describe the statistical models separately since fitting the models required different approaches. For both model types, we fit vital rate models with and without climate covariates. Models with climate effects contain five climate covariates that we chose \emph{a priori} based on previous model selection efforts using these data (Chu et al. *in press*) and expert advice (Lance Vermeire, *personal communication*): "water year" precipitation at \emph{t}-2 (lagppt); April through June precipitation at \emph{t}-1 and \emph{t} (ppt1 and ppt2, respectively) and April through June temperature at \emph{t}-1 and \emph{t} (TmeanSpr1 and TmeanSpr2, respectively), where \emph{t}-1 to *t* is the transition of interest. We also include interactions among same-year climate covariates (e.g., ppt1 $\times$ TmeansSpr1), resulting in a total of seven climate covariates.

We fit all models using a hierarchical Bayesian approach. The models are fully descibed in Appendix A, so here we focus on the main process and the model likelihood. For the likelihood models, $\textbf{y}^X$ is always the relevant vector of observations for vital rate *X* (*X* = *S*, *G*, *R*, or *P*) for survival, growth, recruitment, or population growth). For example, $\textbf{y}^S$ is a vector of 0's and 1's indicating whether a genet survives from *t* to *t+1*, or not. All model parameters are species-specific, but we omit subscripts for species in model descriptions below to reduce visual clutter. For brevity, we only describe models with climate covariates included, but models without climate covariates are exactly the same as the models described below, just without the climate effects.

@@

where $x$ is the log of genet basal area, $\beta_{0,t}$ is a year specific intercept intercept, $\beta_Q$ is the random effect of quadrat group location, $\beta_{s,t}$ is the year-specific slope parameter for size, **z** is a vector of *i* climate covariates specific to year *t*, $\boldsymbol{\beta}_c$ is a vector of fixed climate effects of length *i*, $\beta_{d,1}$ is the effect of intraspecific crowding experienced by the focal genet (*w*), and $\beta_{d,2}$ is a size by crowding ($xw$) interaction effect.

@@

### Model fitting and stastical regularization
####  Model fitting
Our Bayesian approach to fitting the vital rate models required choosing appropriate priors for unknown parameters and deciding which, if any, of those priors should be hierarchical. We decided to fit models where all terms were fit by species. Within a species, we fit yearly size effects and yearly intercepts hierarchically where year-specific coefficients were drawn from global distributions representing the mean size effect and intercept. Quadrat random effects were also fit hierarchically, with quadrat offsets being drawn from distributions with mean zero and a shared variance term (independent Gaussian priors, Appendix A). Climate effects were modeled as independent covariates whose prior distributions were determined by statistical regularization (see **Statistical regularization: Bayesian ridge regression** below).

@@

### Model validation
To test each model's ability to forecast population states we made out-of-sample predictions using leave-one-year-out cross validation. For both levels of modeling and for models with and without climate covariates, we fit the vital rate models using observations from all years except one, and then used those fitted parameters in the population models to perform a one-step-ahead forecast for the year whose observations were withheld from model fitting. Within each observation year, several quadrats were sampled. We made predictions for each observed quadrat in the focal year, initializing each simulation with cover in the quadrat the previous year. Since we were making quadrat-specific predictions, we incorporated the group random effect on the intercept for both models. We repeated this procedure for all 13 observation years, making 100 one-step-ahead forecasts for each quadrat-year combination with parameter uncertainty included via random draw from the MCMC chain as described above. Random year effects were set to zero since year effects cannot be assigned to unobserved years.

@@

### Forecast horizons
An important feature of any forecasting model is the rate at which forecast skill declines as the time between an observation and a forecast increases; the so-called ecological forecast horizon [@Petchey2015].
To assess the forecast horizons of our models, we iniate the model with the population state at some time *t* and make sequential forecasts of the population at times $t+1, t+2, \dots, t+T$ where *T* is the maximum number of years between the initial year and the final year in our observations.
For example, if we initialize the model with percent cover in 1940, we are able to make five forecasts up to the year 1945.
Models are not re-initialized with observations between years.
Thus, in our current example, the model forecast for percent cover in 1941 has a forecast horizon of one year, the forecast in 1942 has a forecast horizon of two years, and so on.
We ran these simulations for all model types (IPM with/without climate; QBM with/without climate) using mean parameter values for all possible initial years.
Then, for a given forecast horizon, we averaged the correlation between forecasts and observations.
Note that these forecasts are all made using in-sample data since we used model fits from the full data set.
Nonetheless, these simulations offer insight into the differences between model forecast horizons.

<!--
### Testing sensitivity to climate covariates
With our fitted and validated models in hand, we ran simulations for each model type (IPM and QBM) under four climate perturbation scenarios: (1) observed climate, (2) precipitation increased by 1%, (3) temperature increased by 1%, and (4) precipitation and temperature increased by 1%. We ran the simulations for 2,500 time steps, enough to estimate equilibrium cover after discarding an initial 500 time steps as burn-in. Each simulation was run under two parameter scenarios: (1) using mean parameter estimates and (2) using randomly drawn parameters from the MCMC chain. We use (1) to detect the overall sensitivity of equilibrium cover to climate, and we use (2) to show the impact of model and parameter uncertainty on forecast precision.

@@

We expected the IPM to produce more accurate and precise forecasts due to either (1) the smaller sample size of the quadrat level data sets compared to the individual level data sets, leading to larger parameter uncertainty for the QBM, or (2) the QBM climate effects being weakly associated with one or more vital rate climate effects at the individual level. To assess the impact of sample size on QBM parameter uncertainty we refit the QBM statistical model (Eqs. 9-10) after removing sets of 2, 5, 10, and 15 quadrats. We fit 10 models at each level of quadrat removal (2, 5, 10, 15 quadrats), removing a different randomly selected set of quadrats for each fit. We calculated the standard deviation of climate main effects (pptLag, ppt1, ppt2, TmeanSpr1, and TmeanSpr2) for each model and averaged those over replicates within each set of quadrat removals. This allowed us to regress parameter uncertainty against sample size.

@@

Results
-------
Both the IPM and QBM generated skillful one-step-ahead forecasts for out-of-sample observations, with an average correlation between predictions and observations ($\rho$) of 0.73 across all models and species (Fig. 2).
Without climate covariates, the accuracy of forecasts from the IPM were not statistically greater than the accuracy of forecasts from the QBM (Fig. 2) and overall error was similar (mean absolute error; Fig. S1).
With climate covariates, the best out-of-sample predictive model (highest *lppd*) for each species and vital rate typically resulted from highly constrained priors on the climate effects (Fig. S2).
Thus, the posterior distributions of climate effects included in our models overlapped zero and generally shrunk toward zero, though for some species-vital rate combinations important effects (80% credible interval does not include zero) did emerge (Fig. 3). 
Despite the small effects, including climate covariates did increase the accuracy of forecasts for all species except *P. smithii* (Fig. 2).
Increases in accuracy due to the inclusion of climate covariates were not significant (*P* > 0.05 for all comparisons of $\rho$ between climate and no-climate forecasts within model types; Fig. 2).
However, the IPM with climate covariates for *P. secunda* did have significantly lower error than the IPM without climate covariates ($t_{(196)}$ = -1.84, *P* = 0.033; Fig. S2).
In only one case were IPM forecasts significantly more accurate than the QBM (Fig. 2): forecast accuracy of *P. secunda* percent cover from an IPM with climate covariates was greater than the accuracy from an equivalent QBM ($t_{(195)}$ = 1.72, *P* = 0.043).
Forecasts from IPMs with climate covariates had significantly lower error than equivalent QBM forecasts for *P. smithii* ($t_{(216)}$ = -3.49, *P* < 0.001) and *P. secunda* ($t_{(196)}$ = -1.83, *P* = 0.034) (Fig. S2).
Results from all pairwise statistical tests are shown in Table S1.
The accuracy of both model's forecasts declined as the forecast horizon increased, but they did so at similar rates (Fig. 4).
The only exception is for *P. secunda* where forecast accuracy appears to slightly increase with forecast horizon, after an initial decrease from a forecast horizon of one year (Fig. 4).

>>>>

**1.** Rapid environmental change has generated growing interest in forecasts of future population trajectories.
Traditional population models built with detailed demographic observations from one study site can address the impacts of environmental change at particular locations, but are difficult to scale up to the landscape and regional scales relevant to management decisions.

**2.** An alternative is to build models using population-level data which are much easier to collect than individual-level data over broad spatial scales.
However, it is unknown whether models built using population-level data adequately capture the effects of density-dependence and environmental forcing that are necessary to generate skillful forecasts.

**3.** Here, we test the consequences of aggregating individual responses when forecasting the population states and trajectories of four perennial grass species in a semi-arid grassland in Montana, USA.
We parameterized two population models for each species, one based on individual-level data (survival, growth and recruitment) and one on population-level data (percent cover), and compared their forecasting skill and forecast horizons with and without the inclusion of climate covariates.
For both models we used Bayesian ridge regression to weight the influence of climate covariates for optimal prediction.

**4.** In the absence of climate effects, we found no significant difference between the forecasting skill of models based on individual-level data and models based on population-level data.
Climate effects were weak and only marginally increased forecasting skill.
Increases in skill with climate covariates were similar between model types for three of the four species. For the fourth species, forecast accuracy of the individual-level model with climate covariates was significantly higher than the accuracy of the equivalent population-level model.

**5.** *Synthesis.* For our focal species at this particular location, and using our particular statistical models, demographic data was generally unnecessary to achieve skillful [ARE ANY OF THESE FORECASTS SKILLFUL? FOCUS INSTEAD ON MODEL DIFFERENCES] forecasts, though for certain species forecast skill can be gained by using demographic data linked to climate covariates.
We conclude that models based on aggregated individual-level data offer a practical alternative to data-intensive demographic models when species do not respond strongly to interannual variation in weather, but when modeling species that do respond to climate drivers, demographically-based models can generate more skillful forecasts.

@@

Perhaps the greatest challenge for ecology in the 21st century is to forecast the impacts of environmental change [@Clark2001; @Petchey2015]. Forecasts require sophisticated modeling approaches that fully account for uncertainty and variability in both ecological process and model parameters [@Luo2011a; but see @Perretti2013a]. The increasing statistical sophistication of population models [@Rees2009] makes them promising tools for predicting the impacts of environmental change on species persistence and abundance. But reconciling the scales at which population models are parameterized and the scales at which environmental changes play out remains a challenge [@Clark2010; @Freckleton2011; @Queenborough2011; @Clark2012]. Most population models are built using demographic data from a single study site because tracking the fates of individuals is so difficult. The resulting models cannot be applied to the landscape and regional scales relevant to decision-making without information about how the fitted parameters respond to spatial variation in biotic and abiotic drivers [@Sæther2007]. The limited spatial extent of individual-level demographic datasets constrains our ability to use population models to address applied questions about the consequences of climate change.

<!--
The inability of most population models to address landscape-scale problems may explain why land managers and conservation planners have embraced species distribution models (SDMs) [see @Guisan2005 for a review]. SDMs typically rely on easy-to-collect presence/absence data and remotely-sensed environmental covariates that allow researchers to model large spatial extents [e.g., @Maiorano2013]  and project future changes in occupancy over landscapes and regions. However, the limitations of SDMs are well known [@Pearson2003; @Elith2009; @Araujo2012]. Ideally, researchers would provide managers with landscape-scale population models, combining the extent of SDMs with information about dynamics and species abundances [@Schurr2012; @Merow2014].  
-->

Aggregate measures of population status, rather than individual performance, offer an intriguing alternative for modeling populations [@Clark2004; @Freckleton2011]. Population-level data cannot provide inference about demographic mechanisms, but might be sufficient for modeling future population states, especially since such data are feasible to collect across broad spatial extents [e.g., @Queenborough2011]. The choice between individual and population-level data involves a difficult trade-off: while individual-level data are necessary for mechanistic models, population-level data enable models that can be applied over greater spatial and temporal extents. An open question is how much forecasting skill is lost when we build models based on population rather than individual-level data.

To date, most empirical population modelers have relied on individual-level data, with few attempts to capitalize on population-level measures. An important exception was an effort by @Taylor2004 to model the population growth rate of an invasive species to identify the best strategies for invasion control. They used a "density-structured" model where the state variable is a discrete density state rather than a continuous density measure. Such models do not require individual-level demographic data and can adequately describe population dynamics. Building on @Taylor2004, @Freckleton2011 showed that density-structured models compare well to continuous models in theory, and @Queenborough2011 provide empirical evidence that density-structured models are capable of reproducing population dynamics at landscape spatial scales, even if some precision is lost when compared to fully continuous models. However, previous tests of density-structured models have yet to assess their ability to forecast out-of-sample observations, and they have not included environmental covariates, which are necessary to forecast population responses to climate change.

Addressing climate change questions with models fit to population-level data is potentially problematic.  Population growth (or decline) is the outcome of demographic processes such as survival, growth, and recruitment that occur at the level of individual plants. Climate can affect each demographic process in unique, potentially opposing, ways [@Dalgleish2011]. These unique climate responses may be difficult to resolve in statistical models based on population-level data where demographic processes are not identifiable. Futhermore, models based on aggregated data may reflect short-term effects of one vital rate more than others whose importance may only emerge over the long-term. For example, a one-year change in a plant species' cover or biomass might reflect individual growth or shrinkage, whereas the long-term trajectory of the population might be more influenced by recruitment.  If important climate effects are missed because of the aggregation inherent in in population-level data, then population models built with such data will make uninformative or unreliable forecasts.

Here, we compare the forecasting skill of statistical and population models based on aggregated, population-level data with the skill of models based on individual-level data. We used a demographic dataset that tracks the fates of individual plants from four species over 14 years to build two kinds of single-species population models, traditional models using individual growth, survival, and recruitment data and alternative models based on basal cover. We simulated the models to answer two questions: (1) Can population models fit using aggregated individual-level data (percent cover) adequately capture density dependence to produce forecasts as skillful as those from models fit to demographic data? 
(2) Can population models fit using aggregated data adequately capture the influence of climate on population growth and, in turn, produce forecasts as skillful as those from models fit to demographic data? 
[THIS NEXT SENTENCE SEEMS OUT OF PLACE. CAN YOU MODIFY PREVIOUS PARAGRAPH TO ADDRESS DENSITY DEPENDENCE ALONG WITH CLIMATE EFFECTS? E.G. DENSITY DEPENDENCE MIGHT ACT MOST STRONGLY ON RECRUITMENT WHICH IS HARD TO DETECT IN A COVER BASED MODEL?] Both questions are motivated by the fact that the effects of intraspecific competition (density dependence) and interannual weather variability act at the level of the individual [@Clark2011b], meaning that models based on individual-level processes should generate more skillful forecasts.

Materials and Methods
---------------------
### Study site and data
Our demographic data come from a northern mixed grass prairie at the Fort Keogh Livestock and Range Research Laboratory near Miles City, Montana, USA ($46^{\circ}$ 19' N, $105^{\circ}$ 48' W). The dataset is available on Ecological Archives[^link] [@Anderson2011], and interested readers should refer to the metadata for a complete description. The site is 800 m above sea level and mean annual precipitation (1878-2009) is 334 mm, with most annual precipitation falling from April through September. The community is grass-dominated, and we focused on the four most abundant grass species: \emph{Bouteloua gracilis} (BOGR), \emph{Hesperostipa comata} (HECO), \emph{Pascopyrum smithii} (PASM), and \emph{Poa secunda} (POSE) (Fig. 1). *B. gracilis* is a warm-season perennial grass, whereas *H. comata*, *P. smithii*, and *P. secunda* are cool-season perennial grasses. All species typically begin growth in the early spring, reach maximum growth and flower in early to mid summer (May-June), and disperse seed in mid to late summer (July-September). [LAST SENTENCES GLOSSES OVER LIKELY BIG DIFFERENCES IN PHENOLOGY. JUST DESCRIBE TIMING OF GROWING SEASON INSTEAD, TO SET UP CLIMATE COVARIATES?]

@@

We modeled each grass population based on two levels of data: individual and quadrat. The individual data is the "raw" data. For the quadrat-level, we data we simply sum individual basal cover for each quadrat by species. This is equivalent to a near-perfect census of quadrat percent cover because measurement error at the individual-level is small [@Chu2014]. Based on these two datasets of 13 year-to-year transitions, we can compare population models built using individual-level data and aggregated, quadrat-level data. At the individual level we explicitly model three vital rates: growth, survival, and recruitment. At the quadrat level we model population growth as change in percent cover of quadrats with non-zero cover in year *t* and in year *t-1*, ignoring within-quadrat extirpation and colonization events because they are very rare in our time series ($N=16$ and $N=13$, respectively, across all species). Sample sizes for each species and vital rate model are shown in Table 1.

@@

### Stastical models of vital rates
At both levels of inference (individual and quadrat), the building blocks of our population models are vital rate regressions. For individual-level data, we fit regressions for survival, growth, and recruitment for each species. At the quadrat-level, we fit a single regression model for population growth. We describe the statistical models separately since they required different approaches. For both model types, we fit vital rate models with and without climate covariates. Models with climate effects contain five climate covariates that we chose \emph{a priori} based on previous model selection efforts using these data (Chu et al. *in press*) and expert advice (Lance Vermeire, *personal communication*): "water year" precipitation at \emph{t}-2 (lagppt); April through June precipitation at \emph{t}-1 and \emph{t} (ppt1 and ppt2, respectively) and April through June temperature at \emph{t}-1 and \emph{t} (TmeanSpr1 and TmeanSpr2, respectively), where \emph{t}-1 to *t* is the transition of interest. We also include interactions among same-year climate covariates (e.g., ppt1 $\times$ TmeansSpr1), resulting in a total of seven climate covariates.

We fit all models using a hierarchical Bayesian approach. The models are fully descibed in Appendix A, so here we focus on the main process and the model likelihood. For the likelihood models, $\textbf{y}^X$ is always the relevant vector of observations for vital rate *X* (*X* = *S*, *G*, *R*, or *P*) for survival, growth, recruitment, or population growth). For example, $\textbf{y}^S$ is a vector of 0's and 1's indicating whether a genet survives from *t* to *t+1*, or not. All model parameters are species-specific, but we omit subscripts for species in model descriptions below to reduce visual clutter. For brevity, we only describe models with climate covariates included, but models without climate covariates are simply the models described below with the climate effects removed.

@@

where $x$ is the log of genet basal area, $\beta_{0,t}$ is a year specific intercept intercept, $\beta_Q$ is the random effect of quadrat group location, $\beta_{s,t}$ is the year-specific slope parameter for size, **z** is a vector of *i* climate covariates specific to year *t*, $\boldsymbol{\beta}_c$ is a vector of fixed climate effects of length *i*, $\beta_{d,1}$ is the effect of intraspecific crowding experienced by the focal genet ($w$), and $\beta_{d,2}$ is a size by crowding ($xw$) interaction effect.

YOU NEED TO EXPLAIN HOW THE w's ARE CALCULATED. COPY TEXT FROM ADLER ET AL. 2010 or CHU  & ADLER 2015.

@@

### Model fitting and statistical regularization
####  Model fitting
Our Bayesian approach to fitting the vital rate models required choosing appropriate priors for unknown parameters and deciding which, if any, of those priors should be hierarchical. We decided to fit models where all terms were fit by species. Within a species, we fit yearly size effects and yearly intercepts hierarchically where year-specific coefficients were drawn from global distributions representing the mean size effect and intercept. Quadrat random effects were also fit hierarchically, with quadrat offsets being drawn from distributions with mean zero and a shared variance term (independent Gaussian priors, Appendix A). Climate effects were modeled as independent covariates whose prior distributions were determined by statistical regularization (see **Statistical regularization: Bayesian ridge regression** below).

@@

### Model validation
To test each model's ability to forecast population states, we made out-of-sample predictions using leave-one-year-out cross validation. For both levels of modeling and for models with and without climate covariates, we fit the vital rate models using observations from all years except one, and then used those fitted parameters in the population models to perform a one-step-ahead forecast for the year whose observations were withheld from model fitting. Within each observation year, several quadrats were sampled. We made predictions for each observed quadrat in each focal year, initializing each simulation with cover in the quadrat the previous year. Since we were making quadrat-specific predictions, we incorporated the group random effect on the intercept for both models. We repeated this procedure for all 13 observation years, making 100 one-step-ahead forecasts for each quadrat-year combination with parameter uncertainty included via random draw from the MCMC chain as described above. Random year effects were set to zero since year effects cannot be assigned to unobserved years.

@@

### Forecast horizons
An important feature of any forecasting model is the rate at which forecast skill declines as the time between an observation and a forecast increases; the so-called ecological forecast horizon [@Petchey2015].
To assess the forecast horizons of our models, we iniate the model with the population state at some time *t* and make sequential forecasts of the population at times $t+1, t+2, \dots, t+T$ where *T* is the maximum number of years between the initial year and the final year of our observations.
For example, if we initialize the model with percent cover in 1940, we are able to make five forecasts up to the year 1945.
Models are not re-initialized with observations between years.
Thus, in our current example, the model forecast for percent cover in 1941 has a forecast horizon of one year, the forecast in 1942 has a forecast horizon of two years, and so on.
We ran these simulations for all model types (IPM with/without climate; QBM with/without climate) using mean parameter values for all possible initial years.
Then, for a given forecast horizon, we averaged the correlation between forecasts and observations.
Note that these forecasts are all made using in-sample data since we used model fits from the full data set.
Nonetheless, these simulations offer insight into the differences between model forecast horizons.

<!--
### Testing sensitivity to climate covariates
With our fitted and validated models in hand, we ran simulations for each model type (IPM and QBM) under four climate perturbation scenarios: (1) observed climate, (2) precipitation increased by 1%, (3) temperature increased by 1%, and (4) precipitation and temperature increased by 1%. We ran the simulations for 2,500 time steps, enough to estimate equilibrium cover after discarding an initial 500 time steps as burn-in. EXPLAIN HOW CLIMATE COVARIATES ARE DRAWN AT EACH TIME STEP. AND NO RANDOM YEAR EFFECTS? Each simulation was run under two parameter scenarios: (1) using mean parameter estimates and (2) using randomly drawn parameters from the MCMC chain. We use (1) to detect the overall sensitivity of equilibrium cover to climate, and we use (2) to show the impact of model and parameter uncertainty on forecast precision.

@@

We expected the IPM to produce more accurate and precise forecasts due to either (1) the smaller sample size of the quadrat level data sets compared to the individual level data sets, leading to larger parameter uncertainty for the QBM, or (2) the QBM climate effects being weakly associated with one or more vital rate climate effects at the individual level. To assess the impact of sample size on QBM parameter uncertainty, we refit the QBM statistical model (Eqs. 9-10) after removing sets of 2, 5, 10, and 15 quadrats. We fit 10 models at each level of quadrat removal (2, 5, 10, 15 quadrats), removing a different randomly selected set of quadrats for each fit. We calculated the standard deviation of climate main effects (pptLag, ppt1, ppt2, TmeanSpr1, and TmeanSpr2) for each model and averaged those over replicates within each set of quadrat removals. This allowed us to regress parameter uncertainty against sample size.

@@

Results
-------
Both the IPM and QBM generated skillful one-step-ahead forecasts for out-of-sample observations, with an average correlation between predictions and observations ($\rho$) of 0.73 across all models and species (Fig. 2). IS THIS SKILLFUL FORECASTING OR IS THIS JUST STRONG AUTOCORRELATION IN COVER?
Without climate covariates, the accuracy of forecasts from the IPM were not statistically greater than the accuracy of forecasts from the QBM (Fig. 2) and overall error was similar (mean absolute error; Fig. S1).
With climate covariates, the best out-of-sample predictive model (highest *lppd*) for each species and vital rate typically resulted from highly constrained priors on the climate effects (Fig. S2).
Thus, the posterior distributions of climate effects included in our models overlapped zero and generally shrunk toward zero, though for some species-vital rate combinations important effects (80% credible interval does not include zero) did emerge (Fig. 3).

Despite the weak climate effects, including climate covariates did increase the accuracy of forecasts for all species except *P. smithii* (Fig. 2).
Increases in accuracy due to the inclusion of climate covariates were not significant (*P* > 0.05 for all comparisons of $\rho$ between climate and no-climate forecasts within model types; Fig. 2).
PREVIOUS SENTENCE READS AS INCONSISTENT WITH NEXT SENTENCE.
However, the IPM with climate covariates for *P. secunda* did have significantly lower error than the IPM without climate covariates ($t_{(196)}$ = -1.84, *P* = 0.033; Fig. S2).
In only one case were IPM forecasts significantly more accurate than the QBM (Fig. 2): forecast accuracy of *P. secunda* percent cover from an IPM with climate covariates was greater than the accuracy from the QBM with climate covariates ($t_{(195)}$ = 1.72, *P* = 0.043).
WHAT IS THE DIFFERENCE BETWEEN ACCURACY (PREVIOUS SENTENCE) AND ERROR (NEXT SENTENCE)?
Forecasts from IPMs with climate covariates had significantly lower error than equivalent QBM forecasts for *P. smithii* ($t_{(216)}$ = -3.49, *P* < 0.001) and *P. secunda* ($t_{(196)}$ = -1.83, *P* = 0.034) (Fig. S2).
Results from all pairwise statistical tests are shown in Table S1.

The accuracy of both model's forecasts declined as the forecast horizon increased, but they did so at similar rates (Fig. 4).
The only exception is for *P. secunda* where forecast accuracy appears to slightly increase with forecast horizon, after an initial decrease from a forecast horizon of one year (Fig. 4).

I TRIED TO BREAK THIS PARAGRAPH UP TO MAKE RESULTS EASIER TO FOLLOW

IS THIS NEXT PART OLD?
